{\fontfamily{ptm}\selectfont A Parameter Free Accelerated Proximal Gradient Method Without Restarting }

### 2025-01-24 Notebook LaTEX/R-WAPG/paper3.tex
>w<, latex changes on formatting
>w<, changes

% \input{presets/wang_full.tex}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
\input{presets/julia_lstlisting.tex}

@@

\section{Introduction}
    \subsection{Motivations}
        For simplicity consider minimizing a differentiable convex function $F: \RR^n \rightarrow \RR$. 
        Let $F$ be $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex meaning that there exists $L \in \RR$ such that for all $x \in \RR^n, y\in \RR^n$ it has $\mu\Vert x - y\Vert\le \Vert \nabla F(x) - \nabla F(y)\Vert \le L \Vert x - y \Vert$. 
        \par 
        The Nesterov's acceleration scheme which was originally proposed in 1983 \cite{nesterov_method_1983} is a celebrated first order method for solving the minimization problem. 
        Initialize $x_1 = y_1$ and $t_0 = 1$, the algorithm finds $(x_k)_{k \ge 1}$ for all $k \ge 1$ by: 
        \begin{align}
            & x_{k + 1} = y_k - L^{-1}\nabla F(y_k), 
            \\
            & t_{k + 1} = 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right), 
            \\
            & \theta_{k + 1} = (t_{k} - 1)/t_{k + 1}, 
            \\
            & y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k). 
        \end{align}\label{eqn:example_algorithm}
        If the minimizer $x^*$ exists, then the optimality gap $F(x_k) - F(x^*)$ decreases at a rate of $\mathcal O(1/k^2)$, a faster rate compared to gradient descent which is $\mathcal O(1/k)$. 
        It's consider optimal in some sense and see Chapter 2 of Nesterov's book \cite{nesterov_lectures_2018} for details. 
        However, things are not sunshine, rainbow and happily ever after because the above scheme is associated with the issue that if $F$ is $\mu > 0$ strongly convex, the optimality gap $F(x_k) - F(x^*)$ oscillates and converges sub-linearly, making it worse than gradient descent. See page 9 of Su et al. \cite{su_differential_2016}. 
        \par
        This issue motivated a vast amount of literatures aims at improving, interpreting, and extending this method. 
        Restarting is a popular solution to address the issue of obtaining better convergence rate when the objective function is strongly convex. 
        Beck and Teboulle \cite{beck_fast_2009} mitigate the issue by restarting and showed that it has an $\mathcal O(1/k^2)$ convergence still, but it performs empirically better. 
        See \cite{necoara_linear_2019}\cite{aujol_parameter-free_2024} and references within for recent advancements in restarting accelerated proximal gradient algorithm.

@@

        We relax the traditional choice of the sequence $\theta_k$ in Equation \ref{eqn:example_algorithm} and showed an upper bound of the optimal gap. 

@@

        + \frac{L}{2}\Vert x - y\Vert^2,  \label{eqn:pg_model-func}

@@

        \label{eqn:pp_model-func}

@@

        With $\mathcal M^{L^{-1}}, \widetilde{\mathcal M}^{L^{-1}}$ as given by \eqref{eqn:pg_model-func}, \eqref{eqn:pp_model-func}, 

@@

        Proceeding stepwise, we want to specialize the results of Proposition \ref{prop:wapg-convergence} for the above choice of sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$. 

@@

            \item $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ are valid R-WAPG sequences, so we can use the convergence results. 
            \item $\rho_k \ge 1$ so Lemma \ref{lemma:inverted-fista-seq} (iii) applies. 
            \item Simplifies the conclusion of Proposition \ref{prop:wapg-convergence} by the previous two items. 

>>>>

% \input{presets/wang_full.tex}

\usepackage{ifthen}\newboolean{draftmode}\setboolean{draftmode}{true}
\input{presets/wang/input.tex}
\input{presets/misc.tex}
\input{presets/julia_lstlisting.tex}

@@

\section{Introduction}
    \subsection{Motivations}
        For simplicity consider minimizing a differentiable convex function $F: \RR^n \rightarrow \RR$. 
        Let $F$ be $L$-Lipschitz smooth and $\mu \ge 0$ strongly convex meaning that there exists $L \in \RR$ such that for all $x \in \RR^n, y\in \RR^n$ it has $\mu\Vert x - y\Vert\le \Vert \nabla F(x) - \nabla F(y)\Vert \le L \Vert x - y \Vert$. 
        \par 
        The Nesterov's acceleration scheme which was originally proposed in 1983 \cite{nesterov_method_1983} is a celebrated first order method for solving the minimization problem. 
        Initialize $x_1 = y_1$ and $t_0 = 1$, the algorithm finds $(x_k)_{k \ge 1}$ for all $k \ge 1$ by: 
        \begin{align}
            & x_{k + 1} = y_k - L^{-1}\nabla F(y_k), 
            \\
            & t_{k + 1} = 1/2\left(1 + \sqrt{1 + 4t_{k}^2}\right), 
            \\
            & \theta_{k + 1} = (t_{k} - 1)/t_{k + 1}, 
            \\
            & y_{k + 1} = x_{k + 1} + \theta_{k + 1}(x_{k + 1} - x_k). 
        \end{align}\label{eqn:example-algorithm}
        If the minimizer $x^*$ exists, then the optimality gap $F(x_k) - F(x^*)$ decreases at a rate of $\mathcal O(1/k^2)$, a faster rate compared to gradient descent which is $\mathcal O(1/k)$. 
        It's consider optimal in some sense and see Chapter 2 of Nesterov's book \cite{nesterov_lectures_2018} for details. 
        However, things are not sunshine, rainbow and happily ever after because the above scheme is associated with the issue that if $F$ is $\mu > 0$ strongly convex, the optimality gap $F(x_k) - F(x^*)$ oscillates and converges sub-linearly, making it worse than gradient descent. See page 9 of Su et al. \cite{su_differential_2016}. 
        \par
        This issue motivated a vast amount of literatures aims at improving, interpreting, and extending this method. 
        Restarting is a popular solution to address the issue of obtaining better convergence rate when the objective function is strongly convex. 
        Beck and Teboulle \cite{beck_fast_2009} mitigate the issue by restarting and showed that it has an $\mathcal O(1/k^2)$ convergence still, but it performs empirically better. 
        See \cite{necoara_linear_2019}\cite{aujol_parameter-free_2024} and references within for recent advancements in restarting accelerated proximal gradient algorithm.

@@

        We relax the traditional choice of the sequence $\theta_k$ in Equation \ref{eqn:example-algorithm} and showed an upper bound of the optimal gap. 

@@

        + \frac{L}{2}\Vert x - y\Vert^2,  \label{eqn:pg-model-func}

@@

        \label{eqn:pp-model-func}

@@

        With $\mathcal M^{L^{-1}}, \widetilde{\mathcal M}^{L^{-1}}$ as given by \eqref{eqn:pg-model-func}, \eqref{eqn:pp-model-func}, 

@@

        Proceeding stepwise, we want to specialize the results of Proposition \ref{prop:wapg-convergence} for the above choice of sequence $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$  by completing the following steps. 

@@

            \item Show that the $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}$ given are valid R-WAPG sequences, so we can use the convergence results. 
            \item Show $\rho_k \ge 1$ so Lemma \ref{lemma:inverted-fista-seq} (iii) applies. 
            \item Simplify the conclusion of Proposition \ref{prop:wapg-convergence} by the previous two items. 
